### 2014-07-23 doc/paper/paper.tex
Skeleton for paper about Møil.

\documentclass[12pt,a4paper]{scrartcl}

%\usepackage{german}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\usepackage{url}
%\urldef{\urlTcpdump}{\url}{http://www.tcpdump.org/}

\usepackage{hyperref}
\hypersetup{colorlinks=false,pdfborder={0 0 0}}

\usepackage{biblatex}
%\usepackage[style=alphabetic]{biblatex}
\ExecuteBibliographyOptions{hyperref=true,language=ngerman,backref=true}
\addbibresource{bibliography.bib}

%\usepackage{lineno}
%\linenumbers

\include{commands}
\include{font-family}

\begin{document}

	\title{Møil: A web-based administration tool for database-backed mail servers with postfix and dovecot}
	%\subtitle{}
	\author{
		Henning Müller $<$\href{mailto:henning@orgizm.net}{henning@orgizm.net}$>$
	}
	\date{}

	\maketitle

	\section*{Introduction and problem}
		% Usage examples:
		%	\cite{rest}
		%	\s after a paragraph for a little space
		%	\begin{description}
		%		\item[\rm Foo]\ \\
		%			Bar
		%	\end{description}

	\section*{Analysis of existing software}

	\subsection*{Criteria}

	\subsection*{The contestants}

	\section*{A new solution: Møil}

	\printbibliography

\end{document}

### 2014-07-23 doc/paper/paper.tex
Use Guard for auto-building (despite the numerous dependencies).
Sketch for paper contents [skip ci].

@@
 
-	\title{Møil: A web-based administration tool for database-backed mail servers with postfix and dovecot}
-	%\subtitle{}
-	\author{
-		Henning Müller $<$\href{mailto:henning@orgizm.net}{henning@orgizm.net}$>$
-	}
+	\title{Møil: A web-based administration tool for database-backed mail servers}
+	\author{Henning Müller $<$\href{mailto:henning@orgizm.net}{henning@orgizm.net}$>$}
 	\date{}
@@
 
-	\section*{Introduction and problem}
 	% Usage examples:
@@
 
-	\section*{Analysis of existing software}
+	\section*{Introduction}
+		For mail system setups (meaning one or more servers running Mail
+		Submission, Delivery and Transfer Agent software and providing a
+		Message Store for users \cite{mail-architecture}), storing data about
+		which domains and E-Mail addresses the system is responsible for in
+		databases is beneficial for several reasons.
+
+		% Reasons for database-backing
+		%	Structural clarity for administrators
+		%	Horizontal scalability
+		%	API (e.g. for accounting)
+		% Query possibilities (LDAP, SQL)
+
+	\section*{Evaluation of existing software}
+		% Overview and inspiration
 
 		\subsection*{Criteria}
+			% Data model
+			%	Structure (postfix.admin compatible?)
+			%	Maintainability of schema
+			%	Security (e.g. password hashing)
+			% Responsiveness
+			%	Terminal
+			%	Desktop web browsers
+			%	Mobile devices
+			% Usability
+			% Scalability (for user exposed functionality)
+			% Features
+			%	Basic domain, mailbox and alias management
+			%	Password change for end users
+			%	Privilege model
+			%	Others
+			%		Audit trail
+			%		Vacation
+			%		Spam frontend
+			%		Sieve
+			%		Webmail integration
+			% Programming language
+			% Containability (e.g. independence of Message Store file system)
+			% Database independence
 
 		\subsection*{The contestants}
+			% PHP
+			%	ratuus http://www.ratuus.org/
+			%	ViMbAdmin http://www.vimbadmin.net/
+			%	postfix.admin http://postfixadmin.sourceforge.net/
+			% Perl
+			%	VBox.Adm http://www.vboxadm.net/
+			% Python
+			%	modoboa http://modoboa.org
 
 	\section*{A new solution: Møil}
+		% Rails
+		% Causes for a new solution
+		%	Data model, migrations
+		%	CLI
+		%	Password hash generation for compatibility to older solution
+		%	Other features
+
+		\subsection*{Features}
+
+		\subsection*{Architecture}
+			% Modules
+
+		\subsection*{Data model}
+			% Data model
+			%	Structure / ERD
+			%	Migrations
+			%	sha512-crypt
+
+		\subsection*{Setup}
+			% Relevant configs for postfix and dovecot
+			% Setup of rails app
+			% Links to README on GitHub
 

### 2014-07-23 doc/paper/paper.tex
Forgot tests as a feature.
Abstract and introduction.

@@
 
-	\title{Møil: A web-based administration tool for database-backed mail servers}
+	\title{Møil: A web-based administration tool for mail servers backed by relational databases}
 	\author{Henning Müller $<$\href{mailto:henning@orgizm.net}{henning@orgizm.net}$>$}
@@
 
-	% Usage examples:
-	%	\cite{rest}
-	%	\s after a paragraph for a little space
-	%	\begin{description}
-	%		\item[\rm Foo]\ \\
-	%			Bar
-	%	\end{description}
+	\section*{Abstract}
+		This document compares existing administration user interface solutions
+		for mail servers backed by relational databases and introduces yet
+		another one, which is called Møil and is written in Ruby utilizing the
+		Ruby on Rails\footnote{\url{http://rubyonrails.org}} framework. It
+		brings the handy possibility of managing the database schema with
+		migrations, a lot of beautiful, responsive CRUD and some other nice
+		features (see page \pageref{moeil-features}). The code is available from
+		GitHub\footnote{\url{https://github.com/nning/moeil}} under the AGPLv3
+		\cite{agpl}.
 
@@
 		Message Store for users \cite{mail-architecture}), storing data about
-		which domains and E-Mail addresses the system is responsible for in
-		databases is beneficial for several reasons.
+		which domains and E-Mail addresses the system is responsible for
+		(further called meta data) in relational databases (as opposed to plain
+		files for example) is beneficial for several reasons.
 
 		% Reasons for database-backing
+		%	Horizontal scalability
+
+		In bigger setups, it offers the possibility for scaling to more than
+		one node (horizontal scalability) to cope with high mail and user
+		loads: Several servers handling incoming mail or users reading their
+		mail act as readers on one or more database servers. The performance of
+		the frequent lookups of domains, mailboxes and aliases (forwardings)
+		for receiving mail or user logins can be increased. (We can not
+		consider another big problem with horizontal scaling of Message Stores,
+		here -- the actual accommodation of E-Mails.)
+
+		%	API (e.g. for accounting)
+
+		Database Management Systems (DBMS) for relational databases usually
+		provide a query interface with the Structured Query Language (SQL)
+		\cite{sql} for editing the data sets. It is widely supported by
+		programming languages and equally widely spoken among programmers, which
+		makes it valuable for integration of several services dealing with this
+		mail system meta data. For example this can be used to integrate
+		accounting solutions for charging purposes into mail systems.
+
 		%	Structural clarity for administrators
-		%	Horizontal scalability
-		%	API (e.g. for accounting)
-		% Query possibilities (LDAP, SQL)
+
+		Another reason for the usage of relational databases as mail system
+		meta data storage is the increased structural clarity for
+		administrators. In a typical SMTP server setup (without a relational
+		database), the relevant data for domains, mailboxes and aliases is
+		distributed to several files. The existence of a domain is stated
+		independently of the existence of a mailbox of the same domain or the
+		password of this mailbox. The data is not interconnected; a mailbox
+		address is usually not checked against the list of domains for
+		consistency. For bigger mail system setups, this meta data
+		configuration inevitably gets unmaintainable.
 
@@
 			% Database independence
+			% Security, robustness, tests
 
@@
 
+		\label{moeil-features}
 		\subsection*{Features}

### 2014-07-24 doc/paper/paper.tex
Started evaluation and added info table for solutions.

@@
 		features (see page \pageref{moeil-features}). The code is available from
-		GitHub\footnote{\url{https://github.com/nning/moeil}} under the AGPLv3
-		\cite{agpl}.
+		GitHub\footnote{\url{https://github.com/nning/moeil}} licensed under
+		AGPLv3 \cite{agpl}.
 
 	\section*{Introduction}
+		% Reasons for database-backing
+
 		For mail system setups (meaning one or more servers running Mail
@@
 
-		% Reasons for database-backing
 		%	Horizontal scalability
@@
 		the frequent lookups of domains, mailboxes and aliases (forwardings)
-		for receiving mail or user logins can be increased. (We can not
-		consider another big problem with horizontal scaling of Message Stores,
-		here -- the actual accommodation of E-Mails.)
+		for receiving mail or user logins can be increased. (The actual
+		accommodation of E-Mails -- another big problem with horizontal scaling
+		-- can not be considered here.)
 
@@
 		% Overview and inspiration
+
+		Although an extensive evaluation of the existing administration user
+		interfaces did not happen before implementing a custom solution, some
+		of the programs were still sighted and some of the following criteria
+		was still considered. This more systematical and detailed analysis
+		shall gain an overview and provide inspiration for the development of
+		Møil.
 
@@
 		\subsection*{The contestants}
-			% PHP
-			%	ratuus http://www.ratuus.org/
-			%	ViMbAdmin http://www.vimbadmin.net/
-			%	postfix.admin http://postfixadmin.sourceforge.net/
-			% Perl
-			%	VBox.Adm http://www.vboxadm.net/
-			% Python
-			%	modoboa http://modoboa.org
+			\subsubsection*{modoboa}
+				\infotable{http://modoboa.org}{Python}
 
-	\section*{A new solution: Møil}
+			\subsubsection*{postfix.admin}
+				\infotable{http://postfixadmin.sourceforge.net}{PHP}
+
+			\subsubsection*{ratuus}
+				\infotable{http://www.ratuus.org}{PHP}
+
+			\subsubsection*{VBox.Adm}
+				\infotable{http://www.vboxadm.net}{Perl}
+
+			\subsubsection*{ViMbAdmin}
+				\infotable{http://www.vimbadmin.net}{PHP}
+
+	\section*{Møil}
 		% Rails

### 2014-07-24 doc/paper/paper.tex
Added ERD.
Acronyms and text on data model criteria.

@@
 
-%\usepackage{german}
 \usepackage[utf8]{inputenc}
 \usepackage{graphicx}
+\usepackage[ampersand]{easylist}
+\usepackage[printonlyused,withpage]{acronym}
 
@@
 \usepackage{biblatex}
-%\usepackage[style=alphabetic]{biblatex}
 \ExecuteBibliographyOptions{hyperref=true,language=ngerman,backref=true}
 \addbibresource{bibliography.bib}
-
-%\usepackage{lineno}
-%\linenumbers
 
@@
 		for mail servers backed by relational databases and introduces yet
-		another one, which is called Møil and is written in Ruby utilizing the
-		Ruby on Rails\footnote{\url{http://rubyonrails.org}} framework. It
-		brings the handy possibility of managing the database schema with
-		migrations, a lot of beautiful, responsive CRUD and some other nice
-		features (see page \pageref{moeil-features}). The code is available from
+		another one, which is called \emph{Møil} and is written in Ruby
+		utilizing the Ruby on Rails\footnote{\url{http://rubyonrails.org}}
+		framework. It brings the handy possibility of managing the database
+		schema with migrations, a lot of beautiful, responsive \ac{CRUD} and
+		some other nice features (see page \pageref{sec:moeil:features}). The
+		code is available from
 		GitHub\footnote{\url{https://github.com/nning/moeil}} licensed under
@@
 
-		Database Management Systems (DBMS) for relational databases usually
-		provide a query interface with the Structured Query Language (SQL)
-		\cite{sql} for editing the data sets. It is widely supported by
-		programming languages and equally widely spoken among programmers, which
-		makes it valuable for integration of several services dealing with this
-		mail system meta data. For example this can be used to integrate
-		accounting solutions for charging purposes into mail systems.
+		\acp{DBMS} for relational databases usually provide a query interface
+		with the \ac{SQL} \cite{sql} for editing the data sets. It is widely
+		supported by programming languages and equally widely spoken among
+		programmers, which makes it valuable for integration of several
+		services dealing with this mail system meta data. For example this can
+		be used to integrate accounting solutions for charging purposes into
+		mail systems.
 
@@
 		meta data storage is the increased structural clarity for
-		administrators. In a typical SMTP server setup (without a relational
-		database), the relevant data for domains, mailboxes and aliases is
-		distributed to several files. The existence of a domain is stated
-		independently of the existence of a mailbox of the same domain or the
-		password of this mailbox. The data is not interconnected; a mailbox
-		address is usually not checked against the list of domains for
-		consistency. For bigger mail system setups, this meta data
-		configuration inevitably gets unmaintainable.
+		administrators. In a typical \acs{SMTP} \cite{smtp} server setup
+		(without a relational database), the relevant data for domains,
+		mailboxes and aliases is distributed to several files. The existence of
+		a domain is stated independently of the existence of a mailbox of the
+		same domain or the password of this mailbox. The data is not
+		interconnected; a mailbox address is usually not checked against the
+		list of domains for consistency. For bigger mail system setups, this
+		meta data configuration inevitably gets unmaintainable.
 
@@
 		shall gain an overview and provide inspiration for the development of
-		Møil.
+		\emph{Møil}.
 
 		\subsection*{Criteria}
-			% Data model
-			%	Structure (postfix.admin compatible?)
-			%	Maintainability of schema
-			%	Security (e.g. password hashing)
-			% Responsiveness
-			%	Terminal
-			%	Desktop web browsers
-			%	Mobile devices
-			% Usability
-			% Scalability (for user exposed functionality)
-			% Features
-			%	Basic domain, mailbox and alias management
-			%	Password change for end users
-			%	Privilege model
-			%	Others
-			%		Audit trail
-			%		Vacation
-			%		Spam frontend
-			%		Sieve
-			%		Webmail integration
-			% Programming language
-			% Containability (e.g. independence of Message Store file system)
-			% Database independence
-			% Security, robustness, tests
+			Not only for the development of \emph{Møil}, some of the criteria
+			arose from shortcomings of \emph{postfix.admin} (see page
+			\pageref{sec:contestants:postfix.admin}), which is a quite widely
+			deployed solution.
+
+			\subsubsection*{Data model (1)}
+				From a database engineering perspective, it is in general
+				favourable to design a database with a \crit{normalized
+				relational database schema}{1.1} \cite{dbnorm}. For a query to
+				check the existence of a mailbox for example, a JOIN or nested
+				query would be necessary, which has a very small performance
+				impact. However, the schema normalization prevents data
+				redundancies and inconsistencies. \emph{postfix.admin} does not
+				have a normalized schema, so any solution which has one can not
+				be \crit{compatible to \emph{postfix.admin}}{1.2}. A
+				compatibility could be useful, if there is a big database
+				already administered with \emph{postfix.admin}, which shall be
+				migrated to another tool. Another possibility for this is an
+				\crit{import of \emph{postfix.admin} data}{1.3}. When it comes
+				to updates to the administration interface, \crit{means to
+				handle database schema updates}{1.4} ease the update process
+				(and therefore increase the probability of the actual
+				conduction of an update). Last but not least, \crit{database
+				independence}{1.5} allows a free choice of the \ac{DBMS} and
+				integration with other software.
+
+			\subsubsection*{Numbered index}
+				\begin{easylist}
+					& Data model
+					&& Normalized schema
+					&& \emph{postfix.admin} compatible
+					&& \emph{postfix.admin} import
+					&& Updatable (Migrations)
+					&& Database independence
+
+					& User interface
+					&& Responsiveness
+					&&&	\ac{CLI} support
+					&&& Mobile devices
+					&& Usability
+
+					& Features
+					&& Basic domain, mailbox and alias management
+					&& Password changing for end users
+					&& Privilege seperation
+					&& Others
+					&&& Audit trail
+					&&& Vacation
+					&&& Spam frontend
+					&&& Sieve
+
+					& Security \& robustness
+					&& Password hashing
+					&& Tests
+
+					& Miscellaneous
+					&& Scalability
+					&& Programming language
+					&& Containability (e.g. independence of Message Store file system)
+				\end{easylist}
 
@@
 
+			\label{sec:contestants:postfix.admin}
 			\subsubsection*{postfix.admin}
 				\infotable{http://postfixadmin.sourceforge.net}{PHP}
+
+				As said, \emph{postfix.admin} seems to be a (negative) example
+				not only for \emph{Møil}. Many solutions aim to overcome its
+				shortcomings. \emph{postfix.admin} uses a non-normalized
+				\cite{dbnorm} database schema, which could spare a JOIN in some
+				database queries but is very prone to inconsistencies. There
+				are no means of handling database schema updates; the schema is
+				imported via a \ac{SQL} file on the initial installation. Also
+				it creates an alias entry for every mailbox (even when the mail
+				is not forwarded). This may have a cause, but is messy anyway.
 
@@
 
-		\label{moeil-features}
+		\label{sec:moeil:features}
 		\subsection*{Features}
@@
 			%	Structure / ERD
+			\includegraphics[width=\textwidth]{images/erd.pdf}
 			%	Migrations
@@
 
+	\include{acronyms}
+
 	\printbibliography

### 2014-07-24 doc/paper/paper.tex
Features criteria text and criteria index in appendix.

@@
 \usepackage[ampersand]{easylist}
-\usepackage[printonlyused,withpage]{acronym}
+\usepackage[printonlyused]{acronym}
 
@@
 			\pageref{sec:contestants:postfix.admin}), which is a quite widely
-			deployed solution.
+			deployed solution. The criteria will be classified into five
+			categories (Data model, features, security \& robustness, user
+			interface and miscellaneous) and numbered for easier reference.
 
@@
 
-			\subsubsection*{Numbered index}
-				\begin{easylist}
-					& Data model
-					&& Normalized schema
-					&& \emph{postfix.admin} compatible
-					&& \emph{postfix.admin} import
-					&& Updatable (Migrations)
-					&& Database independence
+			\subsubsection*{Features (2)}
+				There are features, without which a mail server admistration
+				user interface would not work at all and yet I list them,
+				because it is more about analyzing, how these are implemented.
+				The \crit{basic domain, mailbox and alias management}{2.1} is
+				one of them. The \crit{password changing ability for
+				users}{2.2} is a feature, which is also quite basic and the
+				first one legitimating letting users access the administration
+				frontend. Some solutions also integrate into popular
+				self-hosted web mailer software for password changing. When
+				users are allowed to access the administration frontend,
+				\crit{access control}{2.3} or authorization is needed to
+				selectively restrict the access to resources. It could also be
+				possible to delegate administrative access on domains or
+				resources in general to certain users.
+				\s
 
-					& User interface
-					&& Responsiveness
-					&&&	\ac{CLI} support
-					&&& Mobile devices
-					&& Usability
+				More mature utilities also bring useful non-basic features as
+				an \crit{audit trail}{2.4.1} for example, which can be used to
+				log any changes to the database with time and transacting
+				account. It is also possible, the audit trail provides an undo
+				functionality. In some environments users demand \crit{vacation
+				mailing}{2.4.2} to automatically answer any incoming mail with
+				a pre-defined text when they are vacant. Some solutions serve
+				users with the possibility to activate or deactivate it and
+				edit the pre-defined text. Automatic answers can also be solved
+				more generic with \crit{Sieve}{2.4.3} \cite{sieve}, which is a
+				server-side mail filter interface. The protocol for users to
+				edit the server-side filters is called \emph{ManageSieve}
+				\cite{managesieve}. Another common demand of users of a mail
+				system is a \crit{spam frontend}{2.4.4} to review mails, which
+				got quarantined by spam or virus checking software. (Although in
+				my personal experience this false negative classification does
+				occure almost never with a setup common in the Open Source
+				world consisting of \emph{amavisd-new}, \emph{spamassassin} and
+				\emph{clamav}.)
 
-					& Features
-					&& Basic domain, mailbox and alias management
-					&& Password changing for end users
-					&& Privilege seperation
-					&& Others
-					&&& Audit trail
-					&&& Vacation
-					&&& Spam frontend
-					&&& Sieve
+			\subsubsection*{Security \& robustness (3)}
 
-					& Security \& robustness
-					&& Password hashing
-					&& Tests
+			\subsubsection*{User interface (4)}
 
-					& Miscellaneous
-					&& Scalability
-					&& Programming language
-					&& Containability (e.g. independence of Message Store file system)
-				\end{easylist}
+			\subsubsection*{Miscellaneous (5)}
 
@@
 
-				As said, \emph{postfix.admin} seems to be a (negative) example
-				not only for \emph{Møil}. Many solutions aim to overcome its
-				shortcomings. \emph{postfix.admin} uses a non-normalized
-				\cite{dbnorm} database schema, which could spare a JOIN in some
-				database queries but is very prone to inconsistencies. There
-				are no means of handling database schema updates; the schema is
-				imported via a \ac{SQL} file on the initial installation. Also
-				it creates an alias entry for every mailbox (even when the mail
-				is not forwarded). This may have a cause, but is messy anyway.
+%				As said, \emph{postfix.admin} seems to be a (negative) example
+%				not only for \emph{Møil}. Many solutions aim to overcome its
+%				shortcomings. \emph{postfix.admin} uses a non-normalized
+%				\cite{dbnorm} database schema, which could spare a JOIN in some
+%				database queries but is very prone to inconsistencies. There
+%				are no means of handling database schema updates; the schema is
+%				imported via a \ac{SQL} file on the initial installation. Also
+%				it creates an alias entry for every mailbox (even when the mail
+%				is not forwarded). This may have a cause, but is messy anyway.
 
@@
 
+	\appendix
+
+	\newpage
+	\section*{Numbered index of criteria}
+		\begin{easylist}
+			& Data model
+			&& Normalized schema
+			&& \emph{postfix.admin} compatible
+			&& \emph{postfix.admin} import
+			&& Updatable (Migrations)
+			&& Database independence
+
+			& Features
+			&& Basic domain, mailbox and alias management
+			&& Password changing for users
+			&& Access control
+			&& Others
+			&&& Audit trail
+			&&& Vacation
+			&&& Sieve
+			&&& Spam frontend
+
+			& Security \& robustness
+			&& Password hashing
+			&& Tests
+
+			& User interface
+			&& Responsiveness
+			&&&	\ac{CLI} support
+			&&& Mobile devices
+			&& Usability
+
+			& Miscellaneous
+			&& Scalability
+			&& Programming language
+			&& Containability (e.g. independence of Message Store file system)
+		\end{easylist}
+
 	\include{acronyms}

### 2014-07-24 app/helpers/application_helper.rb
Fixed flash_class helper method.

@@ -12,7 +12,7 @@ module ApplicationHelper
 
   # Convert flash message type to bootstrap class.
   def flash_class(type)
-    ({ notice: :info, alert: :warning, error: :danger }[type] or type).to_s
+    ({notice: :info, alert: :warning, error: :danger}[type.to_sym] || type).to_s
   end
 
   # Link to last visited page.

### 2014-07-25 doc/paper/paper.tex
Began Møil chapter (introdution and features).

@@
 %\urldef{\urlTcpdump}{\url}{http://www.tcpdump.org/}
+\urldef{\urlARSupport}{\url}{http://api.rubyonrails.org/classes/ActiveRecord/Migration.html#class-ActiveRecord::Migration-label-Database+support}
 
@@
 		another one, which is called \emph{Møil} and is written in Ruby
-		utilizing the Ruby on Rails\footnote{\url{http://rubyonrails.org}}
+		utilizing the \ac{Rails}\footnote{\url{http://rubyonrails.org}}
 		framework. It brings the handy possibility of managing the database
-		schema with migrations, a lot of beautiful, responsive \ac{CRUD} and
+		schema with migrations, a lot of beautiful, responsive \acs{CRUD} and
 		some other nice features (see page \pageref{sec:moeil:features}). The
@@
 		% Rails
+
+		\emph{Møil} is written in
+		Ruby\footnote{\url{https://www.ruby-lang.org}} utilizing the popular
+		\ac{Rails} web framework, which has some features that emphasize it
+		against other frameworks or web development approaches. It promotes the
+		development by \acs{REST} \cite{rest} criteria, supports test driven
+		development and implements many well-known software engineering design
+		patterns and methods of agile software development as Active Record,
+		\ac{CoC}, \ac{DRY} and \ac{MVC}. Additionally, the \ac{Rails} community
+		is quite vivid and there is an huge ecosystem of extensions for many
+		recurrent problems.
+		\s
+
 		% Causes for a new solution
@@
 		%	Password hash generation for compatibility to older solution
-		%	Other features
+
+		That fact alone would maybe suffice as an argument for the development
+		for yet another tool but was not as important as some other ideas.
+		\emph{Møil} began as a replacement for the database schema shipped with
+		\emph{postfix.admin} in a private mail system of mine. The schema
+		was intended to be reimplemented normalized \cite{dbnorm} and with
+		a possibility of being updated through schema
+		migrations\footnote{\url{http://en.wikipedia.org/wiki/Schema_migration}}).
+		Also, the \ac{Rails} project was intended to provide the basis for
+		command line utilities to import the existent data and to perform basic
+		administration tasks on domains, mailboxes and aliases. As there has
+		been only the schema in use, not the whole \emph{postfix.admin} web
+		frontend, there was an extra solution to generate password hashes and
+		send them to an administrative contact in place. A replacement for this
+		extra tool at first was the only function of the web interface part of
+		\emph{Møil}. The migration opportunity was also used to change the
+		password hashing algorithm from unsalted \texttt{md5} to
+		\texttt{sha512-crypt}, which is substantially more secure
+		\cite{missing}. (\texttt{bcrypt} was considered but unfortunately
+		rejected, because it is not supported by \texttt{glibc} included in
+		Debian Linux 7.)
+
+		\todo{Sources for algorithms and their comparisons}
+
+		%	Other features?
 
@@
 		\subsection*{Features}
+			As the time of writing this document, \emph{Møil} comes with the
+			following features.
+
+			\begin{description}
+				\item[\rm Tested with postfix and dovecot]\ \\
+					There are some produtive installations of \emph{Møil} in
+					connection with postfix and dovecot but it could possibly
+					also be used with other mailing software which provides the
+					opportunity to configure the \ac{SQL} statements used to
+					request necessary meta data.
+
+				\item[\rm Track database schema with rails migrations]\ \\
+					Through \emph{Active Record} and \ac{Rails} migrations,
+					\emph{Møil} has a robust solution for database change
+					management. This also makes individual changes to the
+					database schema easily maintainable.
+
+				\item[\rm Database agnostic]\ \\
+					\emph{Møil} is tested mostly on SQLite3 and PostgreSQL but
+					also occasionally on MySQL (or MariaDB respectively). In
+					general, it should be supported on any database supported
+					by \emph{Active Record}, which currently are (at least)
+					MySQL, PostgreSQL, SQLite, SQL Server, Sybase, and all
+					supported Oracle databases except
+					DB2\footnote{\urlARSupport}.
+
+				\item[\rm Managing domains, aliases and mailboxes]\ \\
+					\todo
+
+				\item[\rm Permissions and Roles]\ \\
+					...for Mailboxes to e.g. manage a certain domain.
+					\todo
+
+				\item[\rm Responsive user interface]\ \\
+					\emph{Møil} adapts to the environment, it is needed in. The
+					web user interface is responsive and therefore supports
+					both classic desktop view ports and mobile devices like
+					phones or tablets. For friends of the command line, some
+					commands to manage the most important resources are
+					provided, too.
+
+				\item[\rm Audit log]\ \\
+					Any change to resources is logged with time and acting
+					mailbox to support the collaboration of several different
+					administrative users and to implement the possibility of
+					reverting changes.
+
+				\item[\rm Search]\ \\
+					A search for any resources is provided, which is optionally
+					based on \ac{SQL} or elasticsearch for really huge
+					databases (if installed and configured to be used).
+			\end{description}
 
@@
 			%	sha512-crypt
+
+		\subsection*{Perspective}
+			% Better CLI
+			% Spam
+			% Sieve
 

### 2014-07-25 doc/paper/paper.tex
More about Møil (intro, features, data model).

@@
 	\section*{Møil}
+		The development of \emph{Møil} has been begun in May, 2013 as an Open
+		Source project licensed under the terms of the AGPL, Version 3
+		\cite{agpl}.
+
 		% Rails
 
-		\emph{Møil} is written in
-		Ruby\footnote{\url{https://www.ruby-lang.org}} utilizing the popular
-		\ac{Rails} web framework, which has some features that emphasize it
-		against other frameworks or web development approaches. It promotes the
-		development by \acs{REST} \cite{rest} criteria, supports test driven
-		development and implements many well-known software engineering design
-		patterns and methods of agile software development as Active Record,
-		\ac{CoC}, \ac{DRY} and \ac{MVC}. Additionally, the \ac{Rails} community
-		is quite vivid and there is an huge ecosystem of extensions for many
-		recurrent problems.
+		It is written in Ruby\footnote{\url{https://www.ruby-lang.org}}
+		utilizing the popular \ac{Rails} web framework, which has some features
+		that emphasize it against other frameworks or web development
+		approaches. It promotes the development by \acs{REST} \cite{rest}
+		criteria, supports test driven development and implements many
+		well-known software engineering design patterns and methods of agile
+		software development as Active Record, \ac{CoC}, \ac{DRY} and \ac{MVC}.
+		Additionally, the \ac{Rails} community is quite vivid and there is an
+		huge ecosystem of extensions for many recurrent problems.
 		\s
@@
 		\subsection*{Features}
-			As the time of writing this document, \emph{Møil} comes with the
-			following features.
+			At the time of writing of this document, \emph{Møil} comes with the
+			below-mentioned features.
 
 			\begin{description}
+				\item[\rm Managing domains, aliases and mailboxes]\ \\
+					\todo
+
 				\item[\rm Tested with postfix and dovecot]\ \\
@@
 					request necessary meta data.
-
-				\item[\rm Track database schema with rails migrations]\ \\
-					Through \emph{Active Record} and \ac{Rails} migrations,
-					\emph{Møil} has a robust solution for database change
-					management. This also makes individual changes to the
-					database schema easily maintainable.
 
@@
 
-				\item[\rm Managing domains, aliases and mailboxes]\ \\
-					\todo
+				\item[\rm Track database schema with rails migrations]\ \\
+					Through \emph{Active Record} and \ac{Rails} migrations,
+					\emph{Møil} has a robust solution for database change
+					management. This also makes individual changes to the
+					database schema easily maintainable.
 
 				\item[\rm Permissions and Roles]\ \\
-					...for Mailboxes to e.g. manage a certain domain.
-					\todo
+					The abilities to manage domains and/or subordinate resources
+					like mailboxes and aliases can be configured fine grained
+					in order to delegate administrative tasks.
 
@@
 			%	Structure / ERD
+
+			The application model is put into graphs in the following image. It
+			shows the model classes and their relations among each other. (The
+			graph is generated, so unfortunately (for now) the actual attributes
+			saving a relation are missing.)
+
 			\includegraphics[width=\textwidth]{images/erd.pdf}
+
+			The model classes will be examined individually in the following
+			subsections.
+			\s
+
+			In this chapter, whenever a resource type is mentioned, which
+			corresponds to an actual \ac{Rails} model class, it is written in a
+			monospace font and -- as the class name -- capitalized, like
+			\texttt{Mailbox}. Attributes of models are also in monospace font,
+			but lowercase, as the \texttt{username} attribute of a
+			\texttt{Mailbox} for example.
+
+			\subsubsection*{Domains}
+
+			\subsubsection*{Mailboxes}
+
+			\subsubsection*{Aliases}
+
+			\subsubsection*{Permissions}
+				\texttt{Mailbox} records (or users) are also used for the login
+				to \emph{Møil}. There are administrators and regular users.
+				Administrators are allowed to edit any resources. Through a
+				\texttt{Permission}, a \texttt{Mailbox} (the \texttt{subject})
+				can be given access to \texttt{Domains} (the \texttt{item}).
+				The rights on these \texttt{Domains} depend on the
+				\texttt{role} of the user in this context. With an owner
+				\texttt{role}, users can manage any aspect of the
+				\texttt{Domains}. The editor \texttt{role} allows users to
+				manage \texttt{Mailboxes} and \texttt{Aliases} of that.
+				\texttt{Domain}. The \texttt{Permission} model is implemented
+				polymorphic. Its \texttt{subject} and \texttt{item} attributes
+				are not bound to \texttt{Mailboxes} and \texttt{Domains} (as in
+				the current implementation). It is possible to extend the
+				implementation easily to allow even more fine-grained
+				permissions (for certain \texttt{Mailboxes} for example).
+
+			\subsubsection*{Versions}
+				As seen in the graph, \texttt{Domains}, \texttt{Mailboxes},
+				\texttt{Aliases} and \texttt{Permissions} are versioned, which
+				means that changes to resources of any of these types are
+				tracked in the audit log and are revertable for administrators.
+				On any change, a reference to the corresponding resource
+				(\texttt{item\_id} and \texttt{item\_type}), the type of change
+				(\texttt{event}), the transacting individual
+				(\texttt{whodunnit}), the serialized \texttt{object} (only on
+				deletions), the \texttt{object\_changes} (only on updates) and
+				the point in time (\texttt{created\_at}) are saved.
+
+			\subsubsection*{Relocations}
+				The relocation model is in fact updated but actually not
+				further used, currently. Whenever a \texttt{Domain} or
+				\texttt{Mailbox} is renamed, it is necessary to move the
+				corresponding folders in the file system of the Mail Store.
+				These moves are represented as \texttt{Relocation} resources,
+				which can be read by a daemon (yet to be implemented) running
+				on the actual Mail Store.
+				\todo{Other solutions?}
+
 			%	Migrations

### 2014-07-25 app/models/permission.rb
Fixed bug in linking to permissions#edit in versions#index.

@@ -13,8 +13,7 @@ class Permission < ActiveRecord::Base
 
   # Return URL array for editing model instance.
   def edit_url_array
-    [:edit, :admin, item, self] if item
-    nil
+    item ? [:edit, :admin, item, self] : nil
   end
 
   # String representation.

### 2014-07-28 doc/study/study.tex
Little changes to Møil data model and modoboa sections.

--- 2014-07-25 doc/paper/paper.tex
@@
 				\infotable{http://modoboa.org}{Python}
+
+				\emph{modoboa} seems the most mature of the compared tools. It
+				uses the Python Django framework and therefore includes
+				facilities like schema migrations.
 
@@
 				corresponding folders in the file system of the Mail Store.
-				These moves are represented as \texttt{Relocation} resources,
+				These moves are represented as \texttt{Relocation} resources
+				saving \texttt{old\_domain} and \texttt{old\_username} values,
 				which can be read by a daemon (yet to be implemented) running
 				on the actual Mail Store.
+
 				\todo{Other solutions?}

### 2014-07-29 doc/study/study.tex
Also analyze PostVis Admin.
Finished criteria, started comparison (overview almost finished).

@@
 \usepackage[printonlyused]{acronym}
+\usepackage{lscape}
+\usepackage{amssymb}
 
@@
 \urldef{\urlARSupport}{\url}{http://api.rubyonrails.org/classes/ActiveRecord/Migration.html#class-ActiveRecord::Migration-label-Database+support}
+\urldef{\urlPHPBashing}{\url}{http://eev.ee/blog/2012/04/09/php-a-fractal-of-bad-design}
 
@@
 		schema with migrations, a lot of beautiful, responsive \acs{CRUD} and
-		some other nice features (see page \pageref{sec:moeil:features}). The
-		code is available from
+		some other nice features (see \emph{Møil} \textref{sec:moeil:features}).
+		The code is available from
 		GitHub\footnote{\url{https://github.com/nning/moeil}} licensed under
@@
 		shall gain an overview and provide inspiration for the development of
-		\emph{Møil}.
+		\emph{Møil}. An in-depth analysis of the source code can not be
+		provided here. Instead, the solutions will be compared by a composed set
+		of criteria. The current state of the source code was checked out on
+		29\textsuperscript{th} July, 2014 from the particular version control
+		systems and used for evaluation.
 
 		\subsection*{Criteria}
+			\label{sec:evaluation:criteria}
 			Not only for the development of \emph{Møil}, some of the criteria
@@
 			interface and miscellaneous) and numbered for easier reference.
+
+			There is some meta-data like website, programming language, last
+			update in VCS and number of commits, which is not numbered and
+			compared. The used programming language would be a neutral
+			attribute but gets a criterium, because before the development of
+			\emph{Møil}, a decision was made against deploying further PHP
+			applications to accomplish a more maintainable and secure web
+			hosting environment\footnote{\urlPHPBashing}.
 
@@
 			\subsubsection*{Security \& robustness (3)}
+				For now, only three criteria in this category shall be
+				considered, because further security analysis would require an
+				in-depth source code review. The \crit{password hashing}{3.1}
+				availability and algorithm strength is important against
+				unauthorized database access through potential security holes
+				in the application. When the database is maliciously read, the
+				value of the passwords to the attacker is inversely
+				proportional to the time it costs, to crack the password
+				hashing. \crit{Test coverage}{3.2} is important for this
+				evaluation and for security purposes, because it lowers the
+				probability of bugs per lines of code by trend. As
+				\crit{Containability}{3.3}, the modular independence (e.g.
+				independence of the web interface from the Message Store file
+				system) from a security perspective is perceived.
 
-			\subsubsection*{User interface (4)}
-
-			\subsubsection*{Miscellaneous (5)}
+			\subsubsection*{Responsive design (UI) (4)}
+				In the days of rapidly increased mobile device usage, a
+				responsive design (e.g. different layout for different screen
+				sizes) is an important attribute of user interfaces. Usually,
+				responsive design means \crit{adapting to mobile devices}{4.1}
+				as phones and tablet computers. In this document, also the
+				\crit{ability of usage from a \ac{CLI}}{4.2} is considered,
+				because it is widespread among system administrators.
 
 		\subsection*{The contestants}
+			For introduction of "the contestants", the most important
+			attributes and specialities will be listed for each. A more
+			detailed comparison of the features can be found under
+			\textref{sec:evaluation:overview}.
+
 			\subsubsection*{modoboa}
-				\infotable{http://modoboa.org}{Python}
+				\infotable{http://modoboa.org}{Python}{2014-07-28}{1594}
 
@@
 
+				% Criteria
+				% 	2.4.1	No undo.
+				%   4.1		Bad mobile experience.
+
+			\subsubsection*{Møil}
+				\infotable{https://github.com/nning/moeil}{Ruby}{2014-07-26}{353}
+
+				Further information on \emph{Møil} is to be found under
+				\textref{sec:moeil}.
+
+			\subsubsection*{postfix.admin}
 			\label{sec:contestants:postfix.admin}
-			\subsubsection*{postfix.admin}
-				\infotable{http://postfixadmin.sourceforge.net}{PHP}
+				\infotable{http://postfixadmin.sourceforge.net}{PHP}{2014-07-22}{1682}
 
@@
 
+				% Criteria
+				%	1.4		Migrations one big file upgrade.php with functions
+				%			like upgrade_<n>_<db>.
+				%	1.5		No systematic database independence (or object
+				%			relational mapper at all), only manual support for
+				%			mysql and pgsql.
+				%	2.4.1	No undo.
+				%	3.1		Password hash default md5-crypt.
+				%	3.3		Can move mail data if configured and access on Mail
+				%			Store, but disabled by default.
+
+			\subsubsection*{PostVis Admin}
+				\infotable{http://postvisadmin.sourceforge.net}{PHP}{2010-06-24}{183}
+				% SQL injections
+				% Badly strutured
+				% Code duplication
+				% Best example for non-PHP policy
+				%
+				% Criteria
+				%	1.2		Also not normalized like postfix.admin and much
+				%			table names and fields equally named.
+				%	2.3		Even on a quick view, access control does not seem
+				%			effective. Any logged in user (also non-superadmin
+				%			users) seem to be able to delete any other user.
+				%	3.1		Default md5-crypt.
+				%	3.3		Features do not work without shell execution on same
+				%			server (e.g. amavisd-release).
+
 			\subsubsection*{ratuus}
-				\infotable{http://www.ratuus.org}{PHP}
+				\infotable{http://www.ratuus.org}{PHP}{}{}
+
+				The Development website was not linked correctly; the linked
+				hostname did not even exist in DNS, so the URL of the VCS
+				could not be determined and the 0.3-r2 version was examined.
+
+				% Bad struture (almost everything inside include/libratuus.php)
+				% but way more organized (object oriented, systematic 'routing'
+				% and access control) than PostVis Admin, for example. No query
+				% sanitization in mailClass->ask_db(). mysql_query() supports
+				% multiple queries configuration dependently.
+				%
+				% Criteria
+				%	2.4.1	No undo.
+				%	3.1		Only md5-crypt.
 
 			\subsubsection*{VBox.Adm}
-				\infotable{http://www.vboxadm.net}{Perl}
+				\infotable{http://www.vboxadm.net}{Perl}{2012-12-11}{657}
+
+				% Manual SQL queries, but sanitized. Not easy to get
+				% information as for example, if it is database independent.
+				%
+				% Criteria
+				% 	1.5		MySQL dependend in some places (w.g. in maintainance
+				%			code), independend in others (e.g. authentication).
+				%	3.2		Just custom SMTP proxy for spam maintainance
+				%			integration seems to be tested a little.
 
 			\subsubsection*{ViMbAdmin}
-				\infotable{http://www.vimbadmin.net}{PHP}
+				\infotable{http://www.vimbadmin.net}{PHP}{2014-07-23}{337}
+
+				% PHP und ORM!!
+				% Criteria
+				%	2.4.1	No undo.
+				%	2.4.3	Support for access restriction for mailboxes to
+				%			SIEVE.
+				%	3.1		Default is salted MD5, "more" secure algorithms
+				%			only supported with dovecotpw call.
+				%	3.3		See 3.1.
+
+			\begin{landscape}
+				\subsubsection*{Comparison overview}
+				\label{sec:evaluation:overview}
+					% 1. modoboa
+					% 2. Møil
+					% 3. VBox.Adm?
+					% 4. postfix.admin
+					% 5. ratuus
+					% 6. PostVis Admin
+
+					\begin{tabular}{l|cccccc|ccccccccc}
+						              & 1 & 1.1 & 1.2  & 1.3 & 1.4   & 1.5  & 2 & 2.1 & 2.2 & 2.3  & 2.4 & 2.4.1 & 2.4.2 & 2.4.3 & 2.4.4 \\
+						\hline
+						modoboa       &   & \y  &  \n  & \y  &  \y   &  \y  &   & \y  & \y  &  \y  &     & (\y)  & \y    &  \y   & \y    \\
+						Møil          &   & \y  &  \n  & \y  &  \y   &  \y  &   & \y  & \y  &  \y  &     &  \y   & \n    &  \n   & \n    \\
+						postfix.admin &   & \n  &      &     & (\y)  & (\y) &   & \y  & \y  &  \y  &     & (\y)  & \y    &  \n   & \n    \\
+						PostVis Admin &   & \n  & (\y) & \n  &  \n   &  \n  &   & \y  & \n  & (\y) &     &  \n   & \n    &  \n   & \y    \\
+						ratuus        &   & \n  &  \y  & \y  &  \n   &  \n  &   & \y  & \y  &  \y  &     & (\y)  & \n    &  \n   & \n    \\
+						VBox.Adm      &   & \y  &  \n  & \y  &  \n   & (\y) &   & \y  & \y  &  \y  &     &  \n   & \y    &  \n   & \y    \\
+						ViMbAdmin     &   & \y  &  \n  & \n  &       &  \y  &   & \y  & \y  &  \y  &     & (\y)  & \n    & (\n)  & \n    \\
+						\hline
+					\end{tabular}
+					\s
+
+					\noindent
+					\begin{tabular}{l|cccc|ccc|ccc}
+						              & 3 & 3.1  & 3.2  & 3.3  & 4 & 4.1  & 4.2 \\
+						\hline
+						modoboa       &   &  \y  &  \y  &  \y  &   & (\y) & \y  \\
+						Møil          &   &  \y  &  \y  &  \y  &   &  \y  & \y  \\
+						postfix.admin &   & (\y) &  \n  & (\y) &   &  \n  & \y  \\
+						PostVis Admin &   & (\y) &  \n  &  \n  &   &  \n  & \n  \\
+						ratuus        &   & (\y) &  \n  &  \y  &   &  \n  & \n  \\
+						VBox.Adm      &   &  \y  & (\y) &  \y  &   &  \n  & \y  \\
+						ViMbAdmin     &   & (\y) &      & (\n) &   &      &     \\
+						\hline
+					\end{tabular}
+					\s
+
+					(See \textref{sec:evaluation:criteria} and
+					\textref{sec:appendix:criteria}.)
+			\end{landscape}
 
 	\section*{Møil}
+	\label{sec:moeil}
 		The development of \emph{Møil} has been begun in May, 2013 as an Open
@@
 
+		\subsection*{Features}
 		\label{sec:moeil:features}
-		\subsection*{Features}
 			At the time of writing of this document, \emph{Møil} comes with the
@@
 	\section*{Numbered index of criteria}
+	\label{sec:appendix:criteria}
 		\begin{easylist}
@@
 			&& Password hashing
-			&& Tests
+			&& Test coverage
+			&& Containability (e.g. independence of Message Store file system)
 
-			& User interface
-			&& Responsiveness
-			&&&	\ac{CLI} support
-			&&& Mobile devices
-			&& Usability
-
-			& Miscellaneous
-			&& Scalability
-			&& Programming language
-			&& Containability (e.g. independence of Message Store file system)
+			& Responsive design
+			&& Mobile devices
+			&& \ac{CLI} support
 		\end{easylist}

### 2014-07-30 doc/study/study.tex
Finished comparison overview table and summaries for contestants.

@@
 \usepackage{url}
-%\urldef{\urlTcpdump}{\url}{http://www.tcpdump.org/}
 \urldef{\urlARSupport}{\url}{http://api.rubyonrails.org/classes/ActiveRecord/Migration.html#class-ActiveRecord::Migration-label-Database+support}
@@
 
+			\todo{Sources for most of the criteria!}
+
 			\subsubsection*{Data model (1)}
@@
 				(and therefore increase the probability of the actual
-				conduction of an update). Last but not least, \crit{database
-				independence}{1.5} allows a free choice of the \ac{DBMS} and
-				integration with other software.
+				conduction of an update). Last but not least, \crit{systematic
+				database independence}{1.5} abstracts from raw \ac{SQL} queries
+				and allows a free choice of the \ac{DBMS} and integration with
+				other software.
 
@@
 			\subsubsection*{Security \& robustness (3)}
+			\todo{Password hashing algorithms strenghts}
 				For now, only three criteria in this category shall be
@@
 
-			\subsubsection*{Responsive design (UI) (4)}
+			\subsubsection*{Responsive design (\acs{UI}) (4)}
 				In the days of rapidly increased mobile device usage, a
@@
 
-				\emph{modoboa} seems the most mature of the compared tools. It
-				uses the Python Django framework and therefore includes
-				facilities like schema migrations.
-
-				% Criteria
-				% 	2.4.1	No undo.
-				%   4.1		Bad mobile experience.
+				Of the compared tools, \emph{modoboa} seems to be one of the
+				most mature ones. It uses the Python Django framework and
+				therefore includes facilities like schema migrations and an
+				\ac{ORM}. It passes almost every criterium; just when it comes
+				to the \crit{audit trail}{2.4.1}, there is no undo support.
+				Also the user interface on a mobile phone, was a little sloppy.
 
@@
 
-%				As said, \emph{postfix.admin} seems to be a (negative) example
-%				not only for \emph{Møil}. Many solutions aim to overcome its
-%				shortcomings. \emph{postfix.admin} uses a non-normalized
-%				\cite{dbnorm} database schema, which could spare a JOIN in some
-%				database queries but is very prone to inconsistencies. There
-%				are no means of handling database schema updates; the schema is
-%				imported via a \ac{SQL} file on the initial installation. Also
-%				it creates an alias entry for every mailbox (even when the mail
-%				is not forwarded). This may have a cause, but is messy anyway.
-
-				% Criteria
-				%	1.4		Migrations one big file upgrade.php with functions
-				%			like upgrade_<n>_<db>.
-				%	1.5		No systematic database independence (or object
-				%			relational mapper at all), only manual support for
-				%			mysql and pgsql.
-				%	2.4.1	No undo.
-				%	3.1		Password hash default md5-crypt.
-				%	3.3		Can move mail data if configured and access on Mail
-				%			Store, but disabled by default.
+				For many of the solutions, \emph{postfix.admin} seems to be an
+				inspiration and also the impulse for a new implementation. Many
+				solutions aim to overcome its shortcomings, but many also fail
+				to live up to that task (which in this case means every PHP
+				tool). \emph{postfix.admin} uses a non-normalized \cite{dbnorm}
+				database schema, which could spare a JOIN in some database
+				queries but is very prone to inconsistencies. \crit{Schema
+				migrations}{1.4} are not present through a database abstraction
+				library or an \ac{ORM}, but manually implemented rather
+				minimalist. There is one big file (\texttt{upgrade.php})
+				containing a pile of functions like
+				\texttt{upgrade\_<n>\_<db>}, which contain database dependent
+				raw \ac{SQL} queries to perform the transformations on the
+				schema. There is no \crit{systematic database
+				independence}{1.5} or an \ac{ORM}, just "manual" support for
+				MySQL and PostgreSQL. Oddly, for every mailbox there is an
+				alias entry created (even when the mail is not forwarded).
+				\emph{postfix.admin} has support for an \crit{audit
+				trail}{2.4.1} but without undo. The default \crit{password
+				hashing}{3.1} algorithm is \texttt{md5-crypt}. It is
+				\crit{containable}{3.3} by default but does not support the
+				movement of mail data, then, which would require the direct
+				access to the Mail Store.
 
@@
 				\infotable{http://postvisadmin.sourceforge.net}{PHP}{2010-06-24}{183}
-				% SQL injections
-				% Badly strutured
-				% Code duplication
-				% Best example for non-PHP policy
-				%
-				% Criteria
-				%	1.2		Also not normalized like postfix.admin and much
-				%			table names and fields equally named.
-				%	2.3		Even on a quick view, access control does not seem
-				%			effective. Any logged in user (also non-superadmin
-				%			users) seem to be able to delete any other user.
-				%	3.1		Default md5-crypt.
-				%	3.3		Features do not work without shell execution on same
-				%			server (e.g. amavisd-release).
+
+				PostVis Admin is the best acknowledgement for the non-PHP
+				policy it was decided on when searching for a
+				\emph{postfix.admin} replacement. The code is very bad
+				structured and full of redundancies. There is no \crit{schema
+				normalization}{1.1} and the schema seems \crit{compatible to
+				\emph{postifx.admin}}{1.2} but is likely to depart from it in
+				some details. \ac{SQL} injections are findable just by checking
+				for criteria like if there is an \crit{access control}{2.3}
+				feature, which is basically existing but not very effective,
+				because it is not implemented securely. (Any logged in user
+				(also non-superadmin users) seem to be able to delete any other
+				user.) The default \crit{password hashing}{3.1} algorithm is
+				\texttt{md5-crypt} and there are features, which do not work
+				without shell execution privileges on the mail system servers
+				(e.g. \texttt{amavisd-release}, so the solution is not really
+				\crit{containable}{3.3}.
 
@@
 				could not be determined and the 0.3-r2 version was examined.
-
-				% Bad struture (almost everything inside include/libratuus.php)
-				% but way more organized (object oriented, systematic 'routing'
-				% and access control) than PostVis Admin, for example. No query
-				% sanitization in mailClass->ask_db(). mysql_query() supports
-				% multiple queries configuration dependently.
-				%
-				% Criteria
-				%	2.4.1	No undo.
-				%	3.1		Only md5-crypt.
+				The code is not well modularized (almost everything is to be
+				found inside \texttt{include/libratuus.php}), but it is well
+				more organized (when it comes to object orientation, systematic
+				"routing" and access control for example) than \emph{PostVis
+				Admin}. The possibility of \ac{SQL} injections is unclear but
+				there is no query sanitization in the
+				\texttt{mailClass->as\_db()} function, that makes use of
+				\texttt{mysql\_query()}, which in turn supports multiple queries
+				configuration dependently. There is an \crit{audit
+				trail}{2.4.1} feature but it supports no undo. Also the default
+				\crit{password hashing}{3.1} algorithm is \texttt{md5-crypt}.
 
@@
 
-				% Manual SQL queries, but sanitized. Not easy to get
-				% information as for example, if it is database independent.
-				%
-				% Criteria
-				% 	1.5		MySQL dependend in some places (w.g. in maintainance
-				%			code), independend in others (e.g. authentication).
-				%	3.2		Just custom SMTP proxy for spam maintainance
-				%			integration seems to be tested a little.
+				\emph{VBox.Adm} makes use of raw \ac{SQL} queries much, but
+				sanitizes them. At some places (e.g. in a database maintainance
+				script), the queries depend on MySQL. In others (e.g. the
+				authentication), they are abstractly noted and \crit{database
+				independend}{1.5}. As with most other solutions, there are no
+				tests. Solely the custom SMTP proxy for spam maintainance
+				integration seems to be tested a little.
 
@@
 
-				% PHP und ORM!!
-				% Criteria
-				%	2.4.1	No undo.
-				%	2.4.3	Support for access restriction for mailboxes to
-				%			SIEVE.
-				%	3.1		Default is salted MD5, "more" secure algorithms
-				%			only supported with dovecotpw call.
-				%	3.3		See 3.1.
+				The only PHP tool, which uses an \ac{ORM}. (But migrations were
+				not used, yet they are supported by that \ac{ORM}.)
+				\emph{ViMbAdmin} has an \crit{audit log}{2.4.1} available but
+				it does not support undo. There is no real \crit{SIEVE
+				support}{2.4.3}, but an access restriction if configurable for
+				mailboxes on different services like SIEVE. The default
+				\crit{password hashing}{3.1} algorithm is salted \texttt{MD5}
+				(\texttt{md5-crypt}?) and more secure algorithms are only
+				supported with shell execution of \texttt{dovecotpw}, which is
+				bad for \crit{containability}{3.3}. As with many other compared
+				solutions, Twitter bootstrap is used as an \ac{UI} framework,
+				which supports responsive design easily but \emph{ViMbAdmin}
+				seems not to make use of it conciously (there is no menu on
+				mobile phones, for example).
 
@@
 				\label{sec:evaluation:overview}
-					% 1. modoboa
-					% 2. Møil
-					% 3. VBox.Adm?
-					% 4. postfix.admin
-					% 5. ratuus
-					% 6. PostVis Admin
-
 					\begin{tabular}{l|cccccc|ccccccccc}
@@
 						VBox.Adm      &   & \y  &  \n  & \y  &  \n   & (\y) &   & \y  & \y  &  \y  &     &  \n   & \y    &  \n   & \y    \\
-						ViMbAdmin     &   & \y  &  \n  & \n  &       &  \y  &   & \y  & \y  &  \y  &     & (\y)  & \n    & (\n)  & \n    \\
+						ViMbAdmin     &   & \y  &  \n  & \n  &  \n   &  \y  &   & \y  & \y  &  \y  &     & (\y)  & \n    & (\n)  & \n    \\
 						\hline
@@
 						VBox.Adm      &   &  \y  & (\y) &  \y  &   &  \n  & \y  \\
-						ViMbAdmin     &   & (\y) &      & (\n) &   &      &     \\
+						ViMbAdmin     &   & (\y) &  \n  & (\n) &   &  \n  & \n  \\
 						\hline
@@
 		\emph{Møil} began as a replacement for the database schema shipped with
-		\emph{postfix.admin} in a private mail system of mine. The schema
-		was intended to be reimplemented normalized \cite{dbnorm} and with
-		a possibility of being updated through schema
-		migrations\footnote{\url{http://en.wikipedia.org/wiki/Schema_migration}}).
-		Also, the \ac{Rails} project was intended to provide the basis for
-		command line utilities to import the existent data and to perform basic
+		\emph{postfix.admin} in a private mail system of mine. The schema was
+		intended to be reimplemented normalized \cite{dbnorm} and with a
+		possibility of being updated through schema migrations. Also, the
+		\ac{Rails} project was intended to provide the basis for command line
+		utilities to import the existent data and to perform basic
 		administration tasks on domains, mailboxes and aliases. As there has
@@
 		rejected, because it is not supported by \texttt{glibc} included in
-		Debian Linux 7.)
+		Debian Linux 7. Therefore, none of the compared tools supports it
+		natively.)
 
@@
 			&& Updatable (Migrations)
-			&& Database independence
+			&& Systematic database independence
 

### 2014-07-30 doc/study/study.tex
Ran aspell, VCS acronym and text on Møil architecture.

@@
 		This document compares existing administration user interface solutions
-		for mail servers backed by relational databases and introduces yet
-		another one, which is called \emph{Møil} and is written in Ruby
-		utilizing the \ac{Rails}\footnote{\url{http://rubyonrails.org}}
-		framework. It brings the handy possibility of managing the database
-		schema with migrations, a lot of beautiful, responsive \acs{CRUD} and
-		some other nice features (see \emph{Møil} \textref{sec:moeil:features}).
-		The code is available from
-		GitHub\footnote{\url{https://github.com/nning/moeil}} licensed under
-		AGPLv3 \cite{agpl}.
+		for mail servers running \emph{postfix} and backed by relational
+		databases and introduces yet another one, which is called \emph{Møil}
+		and is written in Ruby utilizing the
+		\ac{Rails}\footnote{\url{http://rubyonrails.org}} framework. It brings
+		the handy possibility of managing the database schema with migrations,
+		a lot of beautiful, responsive \acs{CRUD} and some other nice features
+		(see \emph{Møil} \textref{sec:moeil:features}). The code is available
+		from GitHub\footnote{\url{https://github.com/nning/moeil}} licensed
+		under AGPLv3 \cite{agpl}.
 
@@
 		mail act as readers on one or more database servers. The performance of
-		the frequent lookups of domains, mailboxes and aliases (forwardings)
+		the frequent look-up of domains, mailboxes and aliases (forwardings)
 		for receiving mail or user logins can be increased. (The actual
@@
 		\emph{Møil}. An in-depth analysis of the source code can not be
-		provided here. Instead, the solutions will be compared by a composed set
-		of criteria. The current state of the source code was checked out on
-		29\textsuperscript{th} July, 2014 from the particular version control
-		systems and used for evaluation.
+		provided here. Instead, the solutions will be compared by a composed
+		set of criteria. The current state of the source code was checked out
+		on 29\textsuperscript{th} July, 2014 from the particular \acp{VCS} and
+		used for evaluation.
 
@@
 			compared. The used programming language would be a neutral
-			attribute but gets a criterium, because before the development of
+			attribute but gets a criterion, because before the development of
 			\emph{Møil}, a decision was made against deploying further PHP
@@
 			\subsubsection*{Features (2)}
-				There are features, without which a mail server admistration
+				There are features, without which a mail server administration
 				user interface would not work at all and yet I list them,
@@
 				first one legitimating letting users access the administration
-				frontend. Some solutions also integrate into popular
+				front-end. Some solutions also integrate into popular
 				self-hosted web mailer software for password changing. When
-				users are allowed to access the administration frontend,
+				users are allowed to access the administration front-end,
 				\crit{access control}{2.3} or authorization is needed to
@@
 				users with the possibility to activate or deactivate it and
-				edit the pre-defined text. Automatic answers can also be solved
-				more generic with \crit{Sieve}{2.4.3} \cite{sieve}, which is a
+				edit the pretext. Automatic answers can also be solved more
+				generic with \crit{Sieve}{2.4.3} \cite{sieve}, which is a
 				server-side mail filter interface. The protocol for users to
@@
 				\cite{managesieve}. Another common demand of users of a mail
-				system is a \crit{spam frontend}{2.4.4} to review mails, which
-				got quarantined by spam or virus checking software. (Although in
-				my personal experience this false negative classification does
-				occure almost never with a setup common in the Open Source
+				system is a \crit{spam front-end}{2.4.4} to review mails, which
+				got quarantined by spam or virus checking software. (Although
+				in my personal experience this false negative classification
+				does occur almost never with a setup common in the Open Source
 				world consisting of \emph{amavisd-new}, \emph{spamassassin} and
@@
 			\subsubsection*{Security \& robustness (3)}
-			\todo{Password hashing algorithms strenghts}
+			\todo{Password hashing algorithms strengths}
 				For now, only three criteria in this category shall be
@@
 			For introduction of "the contestants", the most important
-			attributes and specialities will be listed for each. A more
-			detailed comparison of the features can be found under
+			attributes and specialties will be listed for each. A more detailed
+			comparison of the features can be found under
 			\textref{sec:evaluation:overview}.
@@
 				therefore includes facilities like schema migrations and an
-				\ac{ORM}. It passes almost every criterium; just when it comes
+				\ac{ORM}. It passes almost every criterion; just when it comes
 				to the \crit{audit trail}{2.4.1}, there is no undo support.
@@
 				normalization}{1.1} and the schema seems \crit{compatible to
-				\emph{postifx.admin}}{1.2} but is likely to depart from it in
+				\emph{postfix.admin}}{1.2} but is likely to depart from it in
 				some details. \ac{SQL} injections are findable just by checking
@@
 				The Development website was not linked correctly; the linked
-				hostname did not even exist in DNS, so the URL of the VCS
+				hostname did not even exist in DNS, so the URL of the \ac{VCS}
 				could not be determined and the 0.3-r2 version was examined.
@@
 				\texttt{mailClass->as\_db()} function, that makes use of
-				\texttt{mysql\_query()}, which in turn supports multiple queries
-				configuration dependently. There is an \crit{audit
+				\texttt{mysql\_query()}, which in turn supports multiple
+				queries configuration dependently. There is an \crit{audit
 				trail}{2.4.1} feature but it supports no undo. Also the default
@@
 				\emph{VBox.Adm} makes use of raw \ac{SQL} queries much, but
-				sanitizes them. At some places (e.g. in a database maintainance
+				sanitizes them. At some places (e.g. in a database maintenance
 				script), the queries depend on MySQL. In others (e.g. the
 				authentication), they are abstractly noted and \crit{database
-				independend}{1.5}. As with most other solutions, there are no
-				tests. Solely the custom SMTP proxy for spam maintainance
+				independent}{1.5}. As with most other solutions, there are no
+				tests. Solely the custom SMTP proxy for spam maintenance
 				integration seems to be tested a little.
@@
 				which supports responsive design easily but \emph{ViMbAdmin}
-				seems not to make use of it conciously (there is no menu on
+				seems not to make use of it consciously (there is no menu on
 				mobile phones, for example).
@@
 		\emph{postfix.admin} in a private mail system of mine. The schema was
-		intended to be reimplemented normalized \cite{dbnorm} and with a
+		intended to be re-implemented normalized \cite{dbnorm} and with a
 		possibility of being updated through schema migrations. Also, the
@@
 		been only the schema in use, not the whole \emph{postfix.admin} web
-		frontend, there was an extra solution to generate password hashes and
+		front-end, there was an extra solution to generate password hashes and
 		send them to an administrative contact in place. A replacement for this
@@
 				\item[\rm Tested with postfix and dovecot]\ \\
-					There are some produtive installations of \emph{Møil} in
+					There are some productive installations of \emph{Møil} in
 					connection with postfix and dovecot but it could possibly
@@
 				\item[\rm Permissions and Roles]\ \\
-					The abilities to manage domains and/or subordinate resources
-					like mailboxes and aliases can be configured fine grained
-					in order to delegate administrative tasks.
+					The abilities to manage domains and/or subordinate
+					resources like mailboxes and aliases can be configured fine
+					grained in order to delegate administrative tasks.
 
@@
 					A search for any resources is provided, which is optionally
-					based on \ac{SQL} or elasticsearch for really huge
-					databases (if installed and configured to be used).
+					based on \ac{SQL} or
+					\emph{Elasticsearch}\footnote{\url{http://www.elasticsearch.org}}
+					for really huge databases (if installed and configured to
+					be used).
 			\end{description}
@@
 			% Modules
+
+			The code is strictly modularized according to the \ac{MVC} pattern,
+			like it is encouraged by \ac{Rails}.
+
+			The model connects to the \ac{DBMS} via TCP/IP or Unix Sockets. The
+			abstractly noted queries are translated to the respective \ac{SQL}
+			(or NoSQL) dialect. All communication with external services like
+			postfix, dovecot or the relocation daemon (see
+			\textref{sec:moeil:model:relocation}) happen solely with the
+			\ac{DBMS} as an intermediary.
+
+			The view templates are written in
+			Haml\footnote{\url{http://haml.info}}, styling is done in SASS and
+			CoffeeScript is used for client-side dynamic functionality. Markup,
+			styling and scripts are delivered to the client after being
+			processed by the \ac{Rails} asset pipeline, which converts HAML to
+			HTML, SASS to CSS and CoffeeScript to JavaScript. It also
+			compresses the output and packs it in as few as possible files to
+			keep the requests to the server low.
+
+			Controllers and their actions are routed to by URL path patterns
+			and are responsible for updating the models' state and providing
+			the according data for the views.
+
+			The search is either solved through using the \ac{DBMS} via
+			\ac{SQL} or by using \emph{Elasticsearch}, which is an indexed
+			search server based on \emph{Apache Lucene}.
 
@@
 			shows the model classes and their relations among each other. (The
-			graph is generated, so unfortunately (for now) the actual attributes
-			saving a relation are missing.)
+			graph is generated, so unfortunately (for now) the actual
+			attributes saving a relation and the cardinalities are missing.)
 
@@
 			\subsubsection*{Relocations}
+			\label{sec:moeil:model:relocation}
 				The relocation model is in fact updated but actually not
@@
 				on the actual Mail Store.
-
-				\todo{Other solutions?}
 
@@
 			&&& Sieve
-			&&& Spam frontend
+			&&& Spam front-end
 

### 2014-07-30 doc/study/study.tex
Missing Møil model descriptions and perspective.

@@
 			\subsubsection*{Domains}
+				For \texttt{Domains} the following attributes are saved per
+				record: \texttt{name} (which is the actual domain- or hostname,
+				\texttt{description}, \texttt{backupmx} (indicates, if the
+				domain is specified as backup MX), \texttt{active},
+				\texttt{quick\_access} (indicates, if the domain edit is
+				reachable through a quick access menu in the \ac{UI}),
+				\texttt{mx\_set} (for automatically saving, if the MX record in
+				the DNS points to the correct mail server) and the timestamps
+				of creation and last update. \texttt{Domains} are connected to
+				many \texttt{Mailboxes} and \texttt{Aliases} to reflect the
+				affiliation of the latter.
 
 			\subsubsection*{Mailboxes}
+				A \texttt{Mailbox} is connected to its \texttt{Domain} and
+				possibly also to one \texttt{Relocation} and has the attributes
+				\texttt{username} (which is the part in front of the
+				domain in the E-Mail address), \texttt{encrypted\_password}
+				(containing the password hash), \texttt{name} (for a given name
+				for example), \texttt{mail\_location} (which is empty by
+				default but can be used to move the users mail data location),
+				\texttt{quota}, \texttt{active}, \texttt{admin} and timestamp
+				attributes for saving the date of creation or the last update.
 
 			\subsubsection*{Aliases}
+				\texttt{Aliases} represent mail forwardings and are also
+				connected to their affiliated \texttt{Domain}. They have the
+				attributes \texttt{username}, \texttt{goto} (which is the
+				target of the forwarding and can contain one or more internal
+				or external E-Mail addresses seperated by commas),
+				\texttt{active}, \texttt{description} and timestamps of
+				creation and last update.
 
@@
 		\subsection*{Perspective}
-			% Better CLI
-			% Spam
-			% Sieve
+			Not least through the evaluation of the different solutions for
+			Open Source mail system administration user interfaces, some
+			possible features for Møil become obvious.
+
+			\begin{description}
+				\item[\rm Better \ac{CLI}]\ \\
+					Currently, Møil has seperate \ac{CLI} tools for the most
+					important tasks like managing domains and mailboxes and
+					importing data from an \emph{postfix.admin} installation.
+					This can be expanded into a single \ac{CLI} utility with
+					better support for the features, the web user interface
+					offers.
+
+				\item[\rm Spam and virus check integration]\ \\
+					Some compared solutions integrate \texttt{amavisd-new} by
+					shared database tables. After this being clear, there is
+					not much left to be solved conceptually for a spam and
+					virus check integration and this becomes a question of time.
+
+				\item[\rm Support for server-side filters]\ \\
+					The support for server-side filters in Møil was already
+					examined and begun but at first, a Ruby library for the
+					ManageSieve protocol has to be implemented, because the
+					only one existing is not maintained anymore and became
+					disfunctional.
+			\end{description}
 
 		\subsection*{Setup}
-			% Relevant configs for postfix and dovecot
-			% Setup of rails app
+			\subsubsection*{Relevant configs for postfix and dovecot}
+
+			\subsubsection*{Setup of the rails app}
+
+			\subsubsection*{Further readings}
 				% Links to README on GitHub

### 2014-07-30 doc/study/study.tex
Turned on TOC, little text for Møil setup.

@@
 
-	\section*{Abstract}
+	\tableofcontents
+
+	\section{Abstract}
 		This document compares existing administration user interface solutions
@@
 
-	\section*{Introduction}
+	\section{Introduction}
 		% Reasons for database-backing
@@
 
-	\section*{Evaluation of existing software}
+	\section{Evaluation of existing software}
 		% Overview and inspiration
@@
 
-		\subsection*{Criteria}
+		\subsection{Criteria}
 		\label{sec:evaluation:criteria}
@@
 
-			\subsubsection*{Data model (1)}
+			\subsubsection{Data model (1)}
 				From a database engineering perspective, it is in general
@@
 
-			\subsubsection*{Features (2)}
+			\subsubsection{Features (2)}
 				There are features, without which a mail server administration
@@
 
-			\subsubsection*{Security \& robustness (3)}
+			\subsubsection{Security \& robustness (3)}
 			\todo{Password hashing algorithms strengths}
@@
 
-			\subsubsection*{Responsive design (\acs{UI}) (4)}
+			\subsubsection{Responsive design (\acs{UI}) (4)}
 				In the days of rapidly increased mobile device usage, a
@@
 
-		\subsection*{The contestants}
+		\subsection{The contestants}
 			For introduction of "the contestants", the most important
@@
 
-			\subsubsection*{modoboa}
+			\subsubsection{modoboa}
 				\infotable{http://modoboa.org}{Python}{2014-07-28}{1594}
@@
 
-			\subsubsection*{Møil}
+			\subsubsection{Møil}
 				\infotable{https://github.com/nning/moeil}{Ruby}{2014-07-26}{353}
@@
 
-			\subsubsection*{postfix.admin}
+			\subsubsection{postfix.admin}
 			\label{sec:contestants:postfix.admin}
@@
 
-			\subsubsection*{PostVis Admin}
+			\subsubsection{PostVis Admin}
 				\infotable{http://postvisadmin.sourceforge.net}{PHP}{2010-06-24}{183}
@@
 
-			\subsubsection*{ratuus}
+			\subsubsection{ratuus}
 				\infotable{http://www.ratuus.org}{PHP}{}{}
@@
 
-			\subsubsection*{VBox.Adm}
+			\subsubsection{VBox.Adm}
 				\infotable{http://www.vboxadm.net}{Perl}{2012-12-11}{657}
@@
 
-			\subsubsection*{ViMbAdmin}
+			\subsubsection{ViMbAdmin}
 				\infotable{http://www.vimbadmin.net}{PHP}{2014-07-23}{337}
@@
 			\begin{landscape}
-				\subsubsection*{Comparison overview}
+				\subsubsection{Comparison overview}
 				\label{sec:evaluation:overview}
@@
 
-	\section*{Møil}
+	\section{Møil}
 	\label{sec:moeil}
@@
 
-		\subsection*{Features}
+		\subsection{Features}
 		\label{sec:moeil:features}
@@
 
-		\subsection*{Architecture}
+		\subsection{Architecture}
 			% Modules
@@
 
-		\subsection*{Data model}
+		\subsection{Data model}
 			% Data model
@@
 
-			\subsubsection*{Domains}
+			\subsubsection{Domains}
 				For \texttt{Domains} the following attributes are saved per
@@
 
-			\subsubsection*{Mailboxes}
+			\subsubsection{Mailboxes}
 				A \texttt{Mailbox} is connected to its \texttt{Domain} and
@@
 
-			\subsubsection*{Aliases}
+			\subsubsection{Aliases}
 				\texttt{Aliases} represent mail forwardings and are also
@@
 
-			\subsubsection*{Permissions}
+			\subsubsection{Permissions}
 				\texttt{Mailbox} records (or users) are also used for the login
@@
 
-			\subsubsection*{Versions}
+			\subsubsection{Versions}
 				As seen in the graph, \texttt{Domains}, \texttt{Mailboxes},
@@
 
-			\subsubsection*{Relocations}
+			\subsubsection{Relocations}
 			\label{sec:moeil:model:relocation}
@@
 
-		\subsection*{Perspective}
+		\subsection{Perspective}
 			Not least through the evaluation of the different solutions for
@@
 
-		\subsection*{Setup}
-			\subsubsection*{Relevant configs for postfix and dovecot}
+		\subsection{Setup}
+			Installation instructions and configuration files will not appear
+			itself in this section. Instead, they will be linked to. This
+			should make this document a little more long-lasting, because setup
+			instructions usually are subject to frequent changes.
 
-			\subsubsection*{Setup of the rails app}
+			\subsubsection{Relevant configs for postfix and dovecot}
+				The relevant configuration files for postfix and dovecot
+				(especially regarding to the \ac{SQL} statements compatible to
+				\emph{Møil}) can be found in the source code repository on
+				GitHub in the folders
+				\texttt{doc/postfix}\footnote{\url{https://github.com/nning/moeil/tree/master/doc/postfix}}
+				and
+				\texttt{doc/dovecot}\footnote{\url{https://github.com/nning/moeil/tree/master/doc/dovecot}}.
 
-			\subsubsection*{Further readings}
-				% Links to README on GitHub
+			\subsubsection{Setup of the rails app}
+				The setup of the rails app for testing or deployment on
+				OpenShift is described in the main
+				README\footnote{\url{https://github.com/nning/moeil}} file of
+				\emph{Møil}. A deployment in web servers like \emph{Apache},
+				\emph{nginx} and \emph{lighttpd} is made similar to any other
+				current \ac{Rails} project. (Also it is possible to deploy with
+				\emph{capistrano}.) More detailed and beginner-friendly
+				instructions will be provided in the main README file, later.
 
@@
 	\newpage
-	\section*{Numbered index of criteria}
+	\section{Numbered index of criteria}
 	\label{sec:appendix:criteria}

### 2014-07-30 doc/study/study.tex
Removed last TODOs.

@@
 	\maketitle
+
+	\begin{center}
+		\textit{Updated versions of this document are to be found at \\
+			\url{https://henning.orgizm.net/doc/moeil.pdf}.}
+	\end{center}
 
@@
 
-			\todo{Sources for most of the criteria!}
-
 			\subsubsection{Data model (1)}
@@
 			\subsubsection{Security \& robustness (3)}
-			\todo{Password hashing algorithms strengths}
 				For now, only three criteria in this category shall be
@@
 				proportional to the time it costs, to crack the password
-				hashing. \crit{Test coverage}{3.2} is important for this
-				evaluation and for security purposes, because it lowers the
-				probability of bugs per lines of code by trend. As
-				\crit{Containability}{3.3}, the modular independence (e.g.
-				independence of the web interface from the Message Store file
-				system) from a security perspective is perceived.
+				hashing. \texttt{MD5} and \texttt{SHA1} based methods are
+				considered unsecure, \texttt{SHA256}, \texttt{SHA384},
+				\texttt{SHA512} and \texttt{bcrypt} are
+				recommended\footnote{\url{http://www.keylength.com/en/8/}}.
+				\crit{Test coverage}{3.2} is important for this evaluation and
+				for security purposes, because it lowers the probability of
+				bugs per lines of code by trend. As \crit{Containability}{3.3},
+				the modular independence (e.g. independence of the web
+				interface from the Message Store file system) from a security
+				perspective is perceived.
 
@@
 
-		\todo{Sources for algorithms and their comparisons}
-
 		%	Other features?
@@
 			\begin{description}
-				\item[\rm Managing domains, aliases and mailboxes]\ \\
-					\todo
+				\item[\rm Managing domains, mailbox and aliases]\ \\
+					\emph{Møil} provides beautiful and usable \acp{UI} for
+					management tasks on domains, mailboxes and aliases hosted
+					on a mail system.
 
-				\item[\rm Tested with postfix and dovecot]\ \\
+				\item[\rm Tested with \emph{postfix} and \emph{dovecot}]\ \\
 					There are some productive installations of \emph{Møil} in
-					connection with postfix and dovecot but it could possibly
-					also be used with other mailing software which provides the
-					opportunity to configure the \ac{SQL} statements used to
-					request necessary meta data.
+					connection with \emph{postfix} and \emph{dovecot} but it
+					could possibly also be used with other mailing software
+					which provides the opportunity to configure the \ac{SQL}
+					statements used to request necessary meta data.
 
 				\item[\rm Database agnostic]\ \\
-					\emph{Møil} is tested mostly on SQLite3 and PostgreSQL but
-					also occasionally on MySQL (or MariaDB respectively). In
-					general, it should be supported on any database supported
-					by \emph{Active Record}, which currently are (at least)
-					MySQL, PostgreSQL, SQLite, SQL Server, Sybase, and all
-					supported Oracle databases except
-					DB2\footnote{\urlARSupport}.
+					\emph{Møil} is tested mostly on \emph{SQLite3} and
+					\emph{PostgreSQL} but also occasionally on \emph{MySQL} (or
+					\emph{MariaDB} respectively). In general, it should be
+					supported on any database supported by \emph{Active
+					Record}, which currently are (at least) \emph{MySQL},
+					\emph{PostgreSQL}, \emph{SQLite}, \emph{SQL Server},
+					\emph{Sybase}, and all supported Oracle databases except
+					\emph{DB2}\footnote{\urlARSupport}.
 
@@
 			(or NoSQL) dialect. All communication with external services like
-			postfix, dovecot or the relocation daemon (see
+			\emph{postfix}, \emph{dovecot} or the relocation daemon (see
 			\textref{sec:moeil:model:relocation}) happen solely with the
@@
 
-			\subsubsection{Relevant configs for postfix and dovecot}
-				The relevant configuration files for postfix and dovecot
-				(especially regarding to the \ac{SQL} statements compatible to
-				\emph{Møil}) can be found in the source code repository on
-				GitHub in the folders
+			\subsubsection{Relevant configs for \emph{postfix} and \emph{dovecot}}
+				The relevant configuration files for \emph{postfix} and
+				\emph{dovecot} (especially regarding to the \ac{SQL} statements
+				compatible to \emph{Møil}) can be found in the source code
+				repository on GitHub in the folders
 				\texttt{doc/postfix}\footnote{\url{https://github.com/nning/moeil/tree/master/doc/postfix}}
@@
 
-			\subsubsection{Setup of the rails app}
+			\subsubsection{Setup of the rails application}
 				The setup of the rails app for testing or deployment on
@@
 				\emph{capistrano}.) More detailed and beginner-friendly
-				instructions will be provided in the main README file, later.
+				instructions will be provided in the main README file at a
+				later point.
 
@@
 			&& \emph{postfix.admin} import
-			&& Updatable (Migrations)
+			&& Schema migrations
 			&& Systematic database independence
+			\s
 
@@
 			&&& Spam front-end
+			\s
 
@@
 			&& Test coverage
-			&& Containability (e.g. independence of Message Store file system)
+			&& Containability
+			\s
 

### 2014-07-30 doc/study/study.tex
Another good source regarding hashing algorithms.
Integrated Sepps comments.

@@
 		the frequent look-up of domains, mailboxes and aliases (forwardings)
-		for receiving mail or user logins can be increased. (The actual
+		for receiving mail or user logins can be increased. The actual
 		accommodation of E-Mails -- another big problem with horizontal scaling
-		-- can not be considered here.)
+		-- can not be considered here.
 
@@
 			\pageref{sec:contestants:postfix.admin}), which is a quite widely
-			deployed solution. The criteria will be classified into five
-			categories (Data model, features, security \& robustness, user
-			interface and miscellaneous) and numbered for easier reference.
-
-			There is some meta-data like website, programming language, last
-			update in VCS and number of commits, which is not numbered and
-			compared. The used programming language would be a neutral
-			attribute but gets a criterion, because before the development of
-			\emph{Møil}, a decision was made against deploying further PHP
-			applications to accomplish a more maintainable and secure web
-			hosting environment\footnote{\urlPHPBashing}.
+			deployed solution. The criteria will be classified into four
+			categories (Data model, features, security \& robustness and user
+			interface) and numbered for easier reference. An example for the
+			textual distinguation is the criterion \crit{normalized database
+			schema}{1.1} as it is written here. There is some meta-data like
+			website, programming language, last update in VCS and number of
+			commits, which is not numbered and compared. The used programming
+			language would be a neutral attribute but gets a criterion, because
+			before the development of \emph{Møil}, a decision was made against
+			deploying further PHP applications to accomplish a more
+			maintainable and secure web hosting
+			environment\footnote{\urlPHPBashing}.
 
@@
 				From a database engineering perspective, it is in general
-				favourable to design a database with a \crit{normalized
+				favourable to design a data\-base with a \crit{normalized
 				relational database schema}{1.1} \cite{dbnorm}. For a query to
@@
 			\subsubsection{Features (2)}
-				There are features, without which a mail server administration
-				user interface would not work at all and yet I list them,
-				because it is more about analyzing, how these are implemented.
-				The \crit{basic domain, mailbox and alias management}{2.1} is
-				one of them. The \crit{password changing ability for
-				users}{2.2} is a feature, which is also quite basic and the
-				first one legitimating letting users access the administration
-				front-end. Some solutions also integrate into popular
-				self-hosted web mailer software for password changing. When
-				users are allowed to access the administration front-end,
-				\crit{access control}{2.3} or authorization is needed to
-				selectively restrict the access to resources. It could also be
-				possible to delegate administrative access on domains or
-				resources in general to certain users.
-				\s
+				There are features, which no mail server administration user
+				interface could spare and yet they will be listed here, because
+				it is important to analyze, how these are implemented and how
+				the implementations differ by solution. The \crit{basic
+				domain, mailbox and alias management}{2.1} is one of them. The
+				\crit{password changing ability for users}{2.2} is a feature,
+				which is also quite basic and the first one legitimating
+				letting users access the administration front-end. Some
+				solutions also integrate into popular self-hosted web mailer
+				software for password changing. When users are allowed to
+				access the administration front-end, \crit{access control}{2.3}
+				or authorization is needed to selectively restrict the access
+				to resources. It could also be possible to delegate
+				administrative access on domains or resources in general to
+				certain users.
 
@@
 				\texttt{SHA512} and \texttt{bcrypt} are
-				recommended\footnote{\url{http://www.keylength.com/en/8/}}.
+				recommended\footnote{\url{http://www.keylength.com/en/8/}}\textsuperscript{,}\footnote{\url{http://codahale.com/how-to-safely-store-a-password}}.
 				\crit{Test coverage}{3.2} is important for this evaluation and
@@
 				to the \crit{audit trail}{2.4.1}, there is no undo support.
-				Also the user interface on a mobile phone, was a little sloppy.
+				Also the user interface on a mobile phone is not fully
+				elaborated and has some minor flaws.
 
@@
 
-				Further information on \emph{Møil} is to be found under
-				\textref{sec:moeil}.
+				Further information on \emph{Møil} is to be found in the
+				section \textref{sec:moeil}.
 
@@
 
-				PostVis Admin is the best acknowledgement for the non-PHP
-				policy it was decided on when searching for a
+				\emph{PostVis Admin} is the best acknowledgement for the
+				non-PHP policy it was decided on when searching for a
 				\emph{postfix.admin} replacement. The code is very bad
@@
 		Source project licensed under the terms of the AGPL, Version 3
-		\cite{agpl}.
-
-		% Rails
-
-		It is written in Ruby\footnote{\url{https://www.ruby-lang.org}}
-		utilizing the popular \ac{Rails} web framework, which has some features
-		that emphasize it against other frameworks or web development
-		approaches. It promotes the development by \acs{REST} \cite{rest}
-		criteria, supports test driven development and implements many
-		well-known software engineering design patterns and methods of agile
-		software development as Active Record, \ac{CoC}, \ac{DRY} and \ac{MVC}.
-		Additionally, the \ac{Rails} community is quite vivid and there is an
-		huge ecosystem of extensions for many recurrent problems.
-		\s
+		\cite{agpl}. It is written in
+		Ruby\footnote{\url{https://www.ruby-lang.org}} utilizing the popular
+		\ac{Rails} web framework, which has some features that emphasize it
+		against other frameworks or web development approaches. It promotes the
+		development by \acs{REST} \cite{rest} criteria, supports test driven
+		development and implements many well-known software engineering design
+		patterns and methods of agile software development as Active Record,
+		\ac{CoC}, \ac{DRY} and \ac{MVC}. Additionally, the \ac{Rails}
+		community is quite vivid and there is an huge ecosystem of extensions
+		for many recurrent problems.
 
@@
 			At the time of writing of this document, \emph{Møil} comes with the
-			below-mentioned features.
+			below-mentioned features:
 
@@
 			The code is strictly modularized according to the \ac{MVC} pattern,
-			like it is encouraged by \ac{Rails}.
-
-			The model connects to the \ac{DBMS} via TCP/IP or Unix Sockets. The
-			abstractly noted queries are translated to the respective \ac{SQL}
-			(or NoSQL) dialect. All communication with external services like
-			\emph{postfix}, \emph{dovecot} or the relocation daemon (see
+			like it is encouraged by \ac{Rails}. The model connects to the
+			\ac{DBMS} via TCP/IP or Unix Sockets. The abstractly noted queries
+			are translated to the respective \ac{SQL} (or NoSQL) dialect. All
+			communication with external services like \emph{postfix},
+			\emph{dovecot} or the relocation daemon (see
 			\textref{sec:moeil:model:relocation}) happen solely with the
-			\ac{DBMS} as an intermediary.
-
-			The view templates are written in
+			\ac{DBMS} as an intermediary. The view templates are written in
 			Haml\footnote{\url{http://haml.info}}, styling is done in SASS and
@@
 			compresses the output and packs it in as few as possible files to
-			keep the requests to the server low.
-
-			Controllers and their actions are routed to by URL path patterns
-			and are responsible for updating the models' state and providing
-			the according data for the views.
-
+			keep the requests to the server low. Controllers and their actions
+			are routed to by URL path patterns and are responsible for updating
+			the models' state and providing the according data for the views.
 			The search is either solved through using the \ac{DBMS} via
@@
 			%	Structure / ERD
+			%	Migrations
+			%	sha512-crypt
 
@@
 
+			\noindent
 			The model classes will be examined individually in the following
-			subsections.
-			\s
-
-			In this chapter, whenever a resource type is mentioned, which
-			corresponds to an actual \ac{Rails} model class, it is written in a
-			monospace font and -- as the class name -- capitalized, like
-			\texttt{Mailbox}. Attributes of models are also in monospace font,
-			but lowercase, as the \texttt{username} attribute of a
-			\texttt{Mailbox} for example.
+			subsections. In this chapter, whenever a resource type is
+			mentioned, which corresponds to an actual \ac{Rails} model class,
+			it is written in a monospace font and -- as the class name --
+			capitalized, like \texttt{Mailbox}. Attributes of models are also
+			in monospace font, but lowercase, as the \texttt{username}
+			attribute of a \texttt{Mailbox} for example.
 
@@
 				attributes for saving the date of creation or the last update.
+				As a password hashing algorithm, currently only
+				\texttt{sha512-crypt} is supported.
 
@@
 				(\texttt{whodunnit}), the serialized \texttt{object} (only on
-				deletions), the \texttt{object\_changes} (only on updates) and
-				the point in time (\texttt{created\_at}) are saved.
+				deletions, can be used for undo), the \texttt{object\_changes}
+				(only on updates) and the point in time (\texttt{created\_at})
+				are saved.
 
@@
 				on the actual Mail Store.
-
-			%	Migrations
-			%	sha512-crypt
 
@@
 				compatible to \emph{Møil}) can be found in the source code
-				repository on GitHub in the folders
-				\texttt{doc/postfix}\footnote{\url{https://github.com/nning/moeil/tree/master/doc/postfix}}
-				and
-				\texttt{doc/dovecot}\footnote{\url{https://github.com/nning/moeil/tree/master/doc/dovecot}}.
+				repository on GitHub in the folders \texttt{doc/postfix} and
+				\texttt{doc/dovecot}\footnote{\url{https://github.com/nning/moeil/tree/master/doc}}.
 

### 2014-07-30 doc/study/study.tex
Removed paragraph indentation in contestant overviews.

@@
 
+				\noindent
 				Of the compared tools, \emph{modoboa} seems to be one of the
@@
 
+				\noindent
 				Further information on \emph{Møil} is to be found in the
@@
 
+				\noindent
 				For many of the solutions, \emph{postfix.admin} seems to be an
@@
 
+				\noindent
 				\emph{PostVis Admin} is the best acknowledgement for the
@@
 
+				\noindent
 				The Development website was not linked correctly; the linked
@@
 
+				\noindent
 				\emph{VBox.Adm} makes use of raw \ac{SQL} queries much, but
@@
 
+				\noindent
 				The only PHP tool, which uses an \ac{ORM}. (But migrations were

### 2014-12-09 lib/search/domain.rb
Support JRuby again.

@@ -1,7 +1,9 @@
 # Search for domains.
 module Search::Domain
   # Search either via SQL or indexed (depending on settings or argument).
-  def self.search(haystack, needle, sql: !Settings.elasticsearch)
+  def self.search(haystack, needle, options = {})
+    sql = options[:sql] || !Settings.elasticsearch
+
     if sql
       sql_search(haystack, needle)
     else

### 2014-12-10 doc/study/study.tex
Optional line numbers for reviews.
Include date in title.

@@
 
+\IfFileExists{review.tex}{
+	\include{review}
+}
+
 \begin{document}
@@
 	\author{Henning Müller $<$\href{mailto:henning@orgizm.net}{henning@orgizm.net}$>$}
-	\date{}
+	\date{\today}
 
@@
 		This document compares existing administration user interface solutions
-		for mail servers running \emph{postfix} and backed by relational
-		databases and introduces yet another one, which is called \emph{Møil}
-		and is written in Ruby utilizing the
-		\ac{Rails}\footnote{\url{http://rubyonrails.org}} framework. It brings
-		the handy possibility of managing the database schema with migrations,
-		a lot of beautiful, responsive \acs{CRUD} and some other nice features
-		(see \emph{Møil} \textref{sec:moeil:features}). The code is available
-		from GitHub\footnote{\url{https://github.com/nning/moeil}} licensed
-		under AGPLv3 \cite{agpl}.
+		for mail servers running \emph{postfix} and \emph{dovecot} backed by
+		relational databases and introduces yet another one, called \emph{Møil}
+		and written in Ruby utilizing the
+		\ac{Rails}\footnote{\url{http://rubyonrails.org}} framework. It
+		provides an usable, responsive \ac{UI}, database independence and ACLs.
+		More features are listed in the \textsecref{sec:moeil:features}. The
+		code is available from
+		GitHub\footnote{\url{https://github.com/nning/moeil}} licensed under
+		AGPLv3 \cite{agpl}.
 
@@
 		Message Store for users \cite{mail-architecture}), storing data about
-		which domains and E-Mail addresses the system is responsible for
+		which domains and email addresses the system is responsible for
 		(further called meta data) in relational databases (as opposed to plain
@@
 		the frequent look-up of domains, mailboxes and aliases (forwardings)
-		for receiving mail or user logins can be increased. The actual
-		accommodation of E-Mails -- another big problem with horizontal scaling
-		-- can not be considered here.
+		for receiving mail or user logins can be increased, because it offers
+		more parallel read capacity. The actual accommodation of emails --
+		another big problem with horizontal scaling	-- can not be covered
+		here.
 
@@
 		\acp{DBMS} for relational databases usually provide a query interface
-		with the \ac{SQL} \cite{sql} for editing the data sets. It is widely
+		based on the \ac{SQL} \cite{sql} for editing the data sets. It is widely
 		supported by programming languages and equally widely spoken among
@@
 		be used to integrate accounting solutions for charging purposes into
-		mail systems.
+		mail systems using \ac{SQL} as consistent interface to the application
+		data (regarded as an \acs{API}).
 
@@
 		\label{sec:evaluation:criteria}
-			Not only for the development of \emph{Møil}, some of the criteria
-			arose from shortcomings of \emph{postfix.admin} (see page
+			Some of the criteria arose from shortcomings of
+			\emph{postfix.admin} (see page
 			\pageref{sec:contestants:postfix.admin}), which is a quite widely
-			deployed solution. The criteria will be classified into four
-			categories (Data model, features, security \& robustness and user
-			interface) and numbered for easier reference. An example for the
-			textual distinguation is the criterion \crit{normalized database
-			schema}{1.1} as it is written here. There is some meta-data like
+			deployed solution -- not only for the development of \emph{Møil}.
+			The criteria will be classified into four categories (Data model,
+			features, security \& robustness and user interface) and numbered
+			for easier reference. An example for the textual distinguation is
+			the criterion \crit{normalized database schema}{1.1} as it is
+			written here: A short description in \textbf{bold} font succeeded
+			by a sequential number in parenthesis. There is some meta-data like
 			website, programming language, last update in VCS and number of
@@
 				interface could spare and yet they will be listed here, because
-				it is important to analyze, how these are implemented and how
+				it is important to analyze how these are implemented and how
 				the implementations differ by solution. The \crit{basic
@@
 				log any changes to the database with time and transacting
-				account. It is also possible, the audit trail provides an undo
+				account. It is also possible the audit trail provides an undo
 				functionality. In some environments users demand \crit{vacation
@@
 				value of the passwords to the attacker is inversely
-				proportional to the time it costs, to crack the password
+				proportional to the time it costs to crack the password
 				hashing. \texttt{MD5} and \texttt{SHA1} based methods are
@@
 			For introduction of "the contestants", the most important
-			attributes and specialties will be listed for each. A more detailed
-			comparison of the features can be found under
-			\textref{sec:evaluation:overview}.
+			attributes and specialties will be listed for each. A more direct
+			comparison of the features and a subsumption can be found in the
+			\textsecref{sec:evaluation:overview}.
 
@@
 				to the \crit{audit trail}{2.4.1}, there is no undo support.
-				Also the user interface on a mobile phone is not fully
-				elaborated and has some minor flaws.
+				Also the \ac{UI} on mobile phones is not fully elaborated and
+				has some minor flaws.
 
@@
 				Further information on \emph{Møil} is to be found in the
-				section \textref{sec:moeil}.
+				\textsecref{sec:moeil}.
 
@@
 				\noindent
-				For many of the solutions, \emph{postfix.admin} seems to be an
+				For many of the solutions \emph{postfix.admin} seems to be an
 				inspiration and also the impulse for a new implementation. Many
@@
 				\item[\rm Responsive user interface]\ \\
-					\emph{Møil} adapts to the environment, it is needed in. The
+					\emph{Møil} adapts to the environment it is needed in. The
 					web user interface is responsive and therefore supports
 					both classic desktop view ports and mobile devices like
-					phones or tablets. For friends of the command line, some
+					phones or tablets. For friends of the command line some
 					commands to manage the most important resources are
@@
 				\item[\rm Search]\ \\
-					A search for any resources is provided, which is optionally
-					based on \ac{SQL} or
+					A combined search for any resource is provided, which is
+					optionally based on \ac{SQL} or
 					\emph{Elasticsearch}\footnote{\url{http://www.elasticsearch.org}}
@@
 			The application model is put into graphs in the following image. It
-			shows the model classes and their relations among each other. (The
-			graph is generated, so unfortunately (for now) the actual
-			attributes saving a relation and the cardinalities are missing.)
+			shows the model classes and their relations among each other. The
+			graph is generated, so unfortunately the actual attributes saving a
+			relation and the cardinalities are missing, for now.
 
@@
 				\texttt{username} (which is the part in front of the
-				domain in the E-Mail address), \texttt{encrypted\_password}
+				domain in the email address), \texttt{encrypted\_password}
 				(containing the password hash), \texttt{name} (for a given name
@@
 				target of the forwarding and can contain one or more internal
-				or external E-Mail addresses seperated by commas),
+				or external email addresses seperated by commas),
 				\texttt{active}, \texttt{description} and timestamps of
@@
 					This can be expanded into a single \ac{CLI} utility with
-					better support for the features, the web user interface
+					better support for the features the web user interface
 					offers.
@@
 				\item[\rm Spam and virus check integration]\ \\
-					Some compared solutions integrate \texttt{amavisd-new} by
-					shared database tables. After this being clear, there is
-					not much left to be solved conceptually for a spam and
-					virus check integration and this becomes a question of time.
+					Of the compared solutions, some integrate
+					\texttt{amavisd-new} by shared database tables. After this
+					being clear, there is not much left to be solved
+					conceptually for a spam and virus check integration and
+					this becomes a question of time.
 
@@
 					The support for server-side filters in Møil was already
-					examined and begun but at first, a Ruby library for the
+					examined and begun but at first a Ruby library for the
 					ManageSieve protocol has to be implemented, because the
 					only one existing is not maintained anymore and became
-					disfunctional.
+					dysfunctional.
 			\end{description}
